---
title: "Nextflow"
author: "Nima Rafati"
date: '2022-07-15'
output: html_document
---

```{r, echo = F}
library(knitr)
library(kableExtra)

```


# DSL2  
## Process  
### input
Inputs are `qualifier`  and `name`.
`qualifier` can be treated as variable type.
- val: Access the input value by name in the process script.
- path: Handle the input value as a path, staging the file properly in the execution context.
- env: Use the input value to set an environment variable in the process script.
- stdin: Forward the input value to the process stdin special file.
- tuple: Handle a group of input values having any of the above qualifiers.
- each: Execute the process for each element in the input collection.

#### `val`
```{bash, eval = F}
process basicExample {
  input:
  val x

  "echo process job $x"
}

workflow {
  def num = Channel.of(1,2,3)
  basicExample(num)
}
```

or 
```{bash, eval = F}
process basicExample {
  input:
  val x

  "echo process job $x"
}
workflow {
  Channel.of(1,2,3).into | basicExample
}
```

### file is deprecated  
It is similar to path; check path instead.  

#### `path`  
Path can be a (groovy) list of files or directories. For example, all the fasta files (*.fa) in a directory.
```{bash, eval = F}
process blastThemAll {
  input:
  path query_file # here you are in a way renaming the variable passed to the process similar to function. 

  "blastp -query ${query_file} -db nr"
}

workflow {
  def proteins = Channel.fromPath( '/some/path/*.fa' )
  blastThemAll(proteins)
}
```

You may also pass a specific file name. All the fastq files in a directory, are passed to `blastp`  as `my_query.fa`.  
```{bash, eval = F}
process blastThemAll {
  input query_file, name: 'my_query.fa' #here input file is named as my_query.fa
  # or
  input 'my_query.fa'
  
  "blastp -query my_query.fa -db nr" #here input file is named as my_query.fa
}

workflow {
  def proteins = Channel.fromPath( '/some/path/*.fa' )
  blastThemAll(proteins)
}
```

There might be situations that you want to pass certain number of files to a process and if the number of files is not equal to the number of files you want to pass, the process will not be executed.  
To address this you can use `arity` parameter.  
```{bash, eval = F}
input:
    path('one.txt', arity: '1')         // exactly one file is expected
    path('two_*.txt', arity: '2')      // exactly two files are expected
    path('several_*.txt', arity: '1..*')   // one or more files are expected
```

You can also control the name of the file in `task` directory by `stageAs` parameter.  
```{bash, eval = F}
input:
    path('one.txt', stageAs: 'my_file.txt')
    # or 
    path x, stageAs: 'my_file.txt'

    "my_command --in my_file.txt"

```



You can also pass multiple files and use a variable ($x in this case) to refer to the file name.  
```{bash, eval = F}
input:
  path x, stageAs: 'my_dir/*.txt'

  "my_command --in $x"

```

#### `env`  
`env` can be used to define a variable in the process script. Opposite to `val` which is used to pass a variable across the processes.
```{bash, eval = F}
process envExample {
  input:
  env my_var

  "echo process job $my_var"
}
``` 


#### `stdin`
```{bash, eval = F}
process stdinExample {
  input:
  stdin my_input

  "echo $my_input"
}
``` 

```{bash, eval = F}
process stdinExample {
  input:
  stdin str

  "cat -"
}

workflow {
  Channel.of('hello', 'hola', 'bonjour', 'ciao')
    | map { it + '\n' } #it represent the element in the channel
    | printAll
}
``` 


#### `tuple` 
tuple is used to pass multiple variables to a process.  
```{bash, eval = F}
process tupleExample {
    input:
    tuple val(x), path('input.txt')

    """
    echo "Processing $x"
    cat input.txt > copy
    """
}

workflow {
  Channel.of( [1, 'alpha.txt'], [2, 'beta.txt'], [3, 'delta.txt'] ) | tupleExample
}
``` 
The output will be :
echo "Processing 1"
cat alpha.txt > copy
echo "Processing 2"
cat beta.txt > copy
echo "Processing 3"
cat delta.txt > copy

#### `each`  
Repeating a process for each element in a collection. 
```{bash, eval = F}
process alignSequences {
  input:
  path in1 # is sequences
  each param1 # is methods
  each path(dirs) # is libraries

  """
  tool -in $in1 -mode $param1 -lib $dirs > result
  """
}

workflow {
  in1 = Channel.fromPath('*.txt')
  param1 = ['R', 'S']
  dirs = [ file('dir1'), file('dir2'), file('dir3') ]

  alignSequences(sequences, methods, libraries)
}
``` 

### output
By output you can define the output of a process. Similar to input, output can also have qualifiers.  
```{bash, eval = F}
process myProcess {
  input:
  path input_file

  output:
  val x
  val 'log'
  val "${input_file.baseName}.out"

  script:
  x = input_file.name
  """
  echo $x > log
  cat $input_file > ${input_file.baseName}.out
  """
}

workflow {
  ch_dummy = Channel.fromPath('*.csv').first()
  (var, log, basename) = myProcess(ch_dummy)

  var.view { "ch_var: $it" }
  log.view { "ch_str: $it" }
  basename.view { "ch_exp: $it" }
}

# or
workflow {
  ch_dummy = Channel.fromPath('*.csv').first()

  results = ch_dummy | myProcess

  # Unpacking the results
  results.view { result ->
    def (var, log, basename) = result
    println "ch_var: $var"
    println "ch_str: $log"
    println "ch_exp: $basename"
  }
}
```

### `path`

```{bash, eval = F}
process myProcess {
  input:
  path input_file

  output:
  path "${input_file.baseName}.out"

  script:
  """
  echo "Processing ${input_file}" > ${input_file.baseName}.out
  """
}

workflow{
  input_file = channel.fromPath('*.txt')
  results = input_files | myProcess
  # or
  results = input_files.map { it -> myProcess(it) }
  # or
  # results = input_files.map { myProcess(it) }
  # or
  # results = input_files.map(myProcess)

  results.view { results ->
  def (output_file) = result
  println "output file: $output_file"
  }
  # or
  # results.view {results ->
  # println "output file: £{results}"
  # }
}
```

### `arity`
Similar to input, you can use `arity` to control the number of files to be generated.  
```{bash, eval = F}
output:
    path('one.txt', arity: '1')         // exactly one file is expected
    path('two_*.txt', arity: '2')      // exactly two files are expected
    path('several_*.txt', arity: '1..*')   // one or more files are expected
```

### Multiple outputs
```{bash, eval = F}
process myProcess {
  input:
  path input_file

  output:
  path "${input_file.baseName}.out"
  path "${input_file.baseName}.log"

  script:
  """
  echo "Processing ${input_file}" > ${input_file.baseName}.out
  echo "Log file for ${input_file}" > ${input_file.baseName}.log
  """
}
workflow{
  input_file = channel.fromPath('*.txt')
  results = input_files | myProcess
  results.view { results ->
  def (output_file, log_file) = result
  println "output file: $output_file"
  println "log file: $log_file"
  }
}
```

### Dynamic output file names
To save files dependent on the input file name.  
```{bash, eval = F}
process mapping {
  input:
  val sample
  path seq

  output:
  path "${sample}.sam"

  """
  tool -in $seq > ${sample}.sam
  """
}
```

### `stdout`
To output the result of a process to the stdout.  
```{bash, eval = F}
process myProcess {
  input:
  path input_file

  output:
  stdout 

  script:
  """
  echo "Processing ${input_file}" 
  """
}
```
### `eval`
`eval` is used to save a standard output of a command. 
```{bash, eval = F}
process myProcess {
  input:
  path input_file

  output:
  eval ('date')

  script:
  """
  echo "Processing ${input_file}" > result.txt
  """
}

```

### `tuple`
BY using `tuple` in output you can define multiple outputs as a channel. For example, you can link the output to information in metadata  such as sample_name.

```{bash, eval = F}
process myProcess {
  input:
  path input_file

  output:
  tuple val(input_file), path("${input_file.baseName}.out")

  script:
  """
  echo "Processing ${input_file}" > ${input_file.baseName}.out
  """
}

workflow{
  input_file = channel.fromPath('*.txt')
  results = input_files | myProcess
  results.view { results ->
  def (input_file, output_file) = result
  println "input file: $input_file"
  println "output file: $output_file"
  }
}
```

### Additional output
#### `emit`
By using `emit` you can access the output of a process in the workflow.  
```{bash, eval = F}
process myProcess {
  output:
  path "test.out", emit: output_file
  # or
  # emit output_file

  script:
  """
  echo "this is a test"> $test.out
  """
}

workflow{
  myProcess()
  myProcess.out.output_file.view()
}
```

#### `optional: true | fales`  
It can be used to switch on or off the output. In other words to give the option to return the output or not.  
```{bash, eval = F}
process myProcess {
  output:
  path "test.out", optional: true

  script:
  """
  echo "this is a test"> $test.out
  """
}
```


### When 
`when` is used to control the execution of a process dependent on state of given condition.  
```{bash, eval = F}
process myProcess {
  input:
  path input_file

  output:
  path "${input_file.baseName}.out"

  when:
  input_file.name =~ /Sample_*.txt/

  script:
  """
  echo "Processing ${input_file}" > ${input_file.baseName}.out
  """
}
```

### Directives
Optional settings for a process.  
#### `accelerator`
To specify a hardware.  
```{bash, eval = F}
process myProcess {
  input:
  path input_file

  output:
  path "${input_file.baseName}.out", emit: output_file
  accelerator 4, type: 'nvidia-h100-mega-80gb'

  script:
  """
  tool --in $input_file --out ${input_file.baseName}.out
  """
}
```

#### `beforeScript` or `afterScript`
This directives are used to run custom script before or after the main process.  
```{bash, eval = F}
process myProcess {
  input:
  path input_file

  output:
  path "${input_file.baseName}.out", emit: output_file

  beforeScript:
  """
  source /path/to/my/script.sh
  """

  script:
  """
  echo "Processing ${input_file}" > ${input_file.baseName}.out
  """

  afterScript:
  """
  echo "Process finished"
  """
}
```
### `cache` 
To cache the output of a process.  
```{bash, eval = F}
process myProcess {
  cahce true # or false 
#  cache lenient: true # minimal input just to save metadata such as file name and size. 
#  cache deep: true # file's content is saved in the cache.
    
  script:
  """
  command....
  """
}
```
### `executor`
To specify the executor.  
- slurm
- local
- k8s
- awsbatch
- sge
- pbspro
- lsf
- .....
```{bash, eval = F}
process myProcess {
  executor 'slurm'
  //...
}
```

You can also create profile where you specify cluster and project.   
```{bash, eval = F}
profiles {

    uppmax {
        executor {
            name = 'slurm'
        }
        process {
            scratch = '$SNIC_TMP'
            clusterOptions = ' -A snic2022-22-606  --mail-type=all --mail-user=nimarafati@gmail.com '
        }
    }
}
``` 

### `clusterOptions`
To specify the settings for a cluster.  
```{bash, eval = F}
process myProcess {
  executor 'slurm'
  clusterOptions '-A snic2022-22-606  --mail-type=all' 
  //...
}
```

### `queue` 
It allows you to set the queue for a process when submitting a job when using grid based executer.  
- `short` for short jobs.  
- `long` for long jobs.  
```{bash, eval = F}
process myProcess {
  queue 'short'
  executor 'sge'
  //...
}
```

### `resourceLabel`  
Using custom-name to pair the workflow to the computing resources (AWS Batch,...). This should be linked according to the structure of the cloud system.  

### `scratch`  
To specify the scratch directory, where tasks output are temporarily stored.  
```{bash, eval = F}
process myProcess {
  scratch '/path/to/scratch'
  //...
}
```

### `shell`  
To specify the shell to run the process. You can have it in configuration file.  
```{bash, eval = F}
process {
  shell '/bin/bash', '-euo', 'pipefail'
}
# or
process.shell = ['/bin/bash', '-euo', 'pipefail']
```

### `conda`
To specify a conda environment.
```{bash, eval = F}
process myProcess {
  conda 'tool=0.1.1'
  //...
}
```

### `container`
To specify a container.
```{bash, eval = F}
process myProcess {
  container 'ubuntu:20.04'
  //...
}
```
### `cpu` & `memory` 
To specify the number of CPUs. 
```{bash, eval = F}
process myProcess {
  cpu 4
  memory '8 GB'
  executor 'slurm'

  script:
  """
  command.... --cpu ${task.cpus}
  """
}
```
### `disk` 
To specify the disk space.    
```{bash, eval = F}
process myProcess {
  disk '100 GB'
  executor 'slurm'
  //...
}
```

### `time` 
To specify the time limit for a process.  
```{bash, eval = F}
process myProcess {
  time '8 h'
  //...
}
```



### `resourceLimit`  
To specify the resource limit for a process. 
```{bash, eval = F}
process myProcess {
  resourceLimit 'cpus: 16, memory: 8GB, time: 8.h, disk: 2GB'
  //...
}
``` 

You can also specify the resource limit in the configuration file.
```{bash, eval = F} 
process {
  resourceLimit 'cpus: 16, memory: 8GB, time: 8.h'
}
```



### `debug` 
It is explanatory, to debug the process.  
```{bash, eval = F}
process myProcess {
  debug true
  //...
}
```


### `errorStrategy`
Tracing and handling the error in the workflow  
```{r, eval = T, echo = F}
df <- data.frame(Name = c('terminate', 'finish', 'ignore', 'retry'), Executor = c('Terminates the execution as soon as an error condition is reported. Pending jobs are killed (default)',
'Initiates an orderly pipeline shutdown when an error condition is raised, waiting the completion of any submitted job.',
'Ignores processes execution errors.',
'Re-submit for execution a process returning an error condition.'))
kable(df, longtable = F, booktabs = T)  %>% kable_styling()
```

You can use `maxErrors 5`, `maxRetries = 3`  together with `retry`.  
```{bash, eval = F}
process ignoreAnyError {
  errorStrategy 'retry'
  maxErrors 5

  script:
  <your command string here>
}
```

### `ext`  
To specify user custom settings.  
```{bash, eval = F}
process myProcess {
  ext version: '1.0', arg: '--param1 value1 --param2 value2'
  //...
}
```

Or in the configuration 
```{bash, eval = F}
params.star_version = '2.7.9.a'
params.star_args = '--runThreadN 8 '
```

### `label` 
To specify a label for a process.
```{bash, eval = F}
process myProcess {
  label 'my_label'
  //...
}
```
You can use the same label for multiple processes. Also, you can use multiple labels for a process.

### `maxForks`  
Define how many process instances to perform in parallel.  
```{bash, eval = F}
process doNotParallelizeIt {
  maxForks 1

  '''
  <your script here>
  '''
}
```

### `module`  
To load a module in local machine/server.   
```{bash, eval = F}
process myProcess {
  module 'BioPerl:blast'
      //...
}
```


### `penv` 
To specify parallel environments.  
```{bash, eval = F}
process myProcess {
  cpu 4
  penv 'openmpi'
  executor 'slurm'
  
  //...
}
```

### `puhlishDir`
To specify the directory to publish the output. Other options to set:  
- `enabled` to enable or disable the publishing.  
- `failOnError` to stop the workflow if the publishing fails.  
- `mode` to specify the mode of publishing:  
  - `copy` to copy the files to the publish directory.  
  - `link` to create a symbolic link to the files in the publish directory.  
  - `move` to move the files to the publish directory.  
  - `rellink` to create a relative symbolic link to the files in the publish directory.  
  - `symlink` to create a symbolic link to the files in the publish directory.  
- `path` to specify the path to the publish directory. `publishDir path: '/path/results/'`  
- `pattern` to specify the pattern of the files to publish. `publishDir pattern: '*.txt'`  
- `saveAs` to specify the name of the file in the publish directory. `publishDir saveAs: 'my_file.txt'`   
- 
```{bash, eval = F}
process myProcess {
  publishDir 'results', mode: 'copy', overwrite: true
  //...
}
```


### `tag` 
To specify a tag for a every analysis in process which will be saved in the log file.  The difference with `label` is that `label` is used to label a group of analysis in processes.  
```{bash, eval = F}
process myProcess {
  tag '$sample'  

  input:
  val sample

  script
  """
  echo "Processing $sample"
  """
}
workflow {
  Channel.of('sampleA', 'sampleB', 'sampleC') | myProcess
}
```
This will give output as:

```
Launching `myProcess` [sampleA]
Launching `myProcess` [sampleB]
Launching `myProcess` [sampleC]
```

### Dynamic directives  
#### Setting resources dynamically  
You can instruct the process to dynamially change the settings dependent on the input. In following example the queue is set to `short` if the input file is smaller than 1MB.  
```{bash, eval = F}
process myProcess {
  input:
  path input_file

  queue {input_file.size < 1.MB ? 'short', 'long'}

  script:
  """
  echo "Processing ${input_file} on queue ${task.queue}"
  """
}
```

Or for computing depedenet on the process labels.  

```{bash, eval = F}
process FASTQC {
    tag "$meta.id"
    label 'process_medium'
    //...
}
process TRIMMOMATIC {
    tag "$meta.id"
    label 'process_low'
    //...
}

process {
  withlabel: 'process_medium' {
        cpus   = { check_max( 2     * task.attempt, 'cpus'    ) }
        memory = { check_max( 12.GB * task.attempt, 'memory'  ) }
        time   = { check_max( 4.h   * task.attempt, 'time'    ) }
    }
    withlabel: 'process_low' {
        cpus   = { check_max( 1     * task.attempt, 'cpus'    ) }
        memory = { check_max( 6.GB  * task.attempt, 'memory'  ) }
        time   = { check_max( 2.h   * task.attempt, 'time'    ) }
    }
  }
```

#### Setting retry dynamically  
You can set the retry dynamically dependent on the availability of ther resources (e.g. network congestion,....).    
```{bash, eval = F}
process {
   errorStrategy = { task.exitStatus in [143,137,104,134,139] ? 'retry' : 'finish' }
    maxRetries    = 1
    maxErrors     = '-1'
}
```
- 143 (SIGTERM): Process was terminated (possibly due to a timeout).  
- 137 (SIGKILL): Process was killed (possibly due to exceeding resource limits).  
- 104: Less common, could be application-specific or network-related.  
- 134 (abort): Process aborted due to internal error.  
- 139 (SIGSEGV): Process encountered a segmentation fault.  

## Channel   
There are two channel types:
- Queue channel: A channel that emits a sequence of values. This channel can be created by:  
  - factory methods: `of`, `fromPath`  
  - operators: `map`, `flatMap` 
  - processes: see above     
- Value channel: A channel that emits a single value. This channel can be created by:  
  - factory methods: `value` 
  - operators: `first`, `collect`, `reduce`  

### Channel factories  
#### `channel.empty`  
Creating an empty channel without emitting any value.  

#### `channel.from`  
Creating a channel from a list of values.   
```{bash, eval = F}
ch = Channel.from( 'alpha', 'beta', 'delta' )
ch.subscribe { println "Got: $it" }
```
- `subscribe` a function for the channel printing the emitted values


#### `channel.fromList` 
Creating a channel from a list of values.  
```{bash, eval = F}
ch = Channel.fromList( ['alpha', 'beta', 'delta'] )
ch.view()

# Or

channel.
  fromList( ['alpha', 'beta', 'delta'] )
  .view { println "Got: $it" }
```
Output:  
```
Got: alpha
Got: beta
Got: delta
```

#### `channel.fromPath`  
Emitting file(s) from a path. 
**NOTE: does not check whether the file exists.**  
```{bash, eval = F}
myFileChannel = Channel.fromPath( '/data/file.txt' )
pairFiles = Channel.fromPath( 'data/file_{1,2}.fq' )
expl3 = Channel.fromPath( '/path/*', hidden: true ) # search for hidden files.   

myFileChannel = Channel.fromPath( '/path/*b', type: 'dir' ) # Listing DIRECTORIES ending with b
myFileChannel = Channel.fromPath( '/path/a*', type: 'any' ) # Listing files and directories starting with a  

Channel.fromPath( ['/some/path/*.txt', '/other/path/*.tx'] ) # You can pass multiple formats to channel  
```

\*\* checkes directories for given files too.   

Parametrs:  
- `maxDepth`: Maximum number of directory levels to visit (default: no limit)  
- `followLinks`: When true it follows symbolic links during directories tree traversal, otherwise they are managed as files (default: true)  
- `relative`: When true returned paths are relative to the top-most common directory (default: false)  
- `checkIfExists`: When true throws an exception of the specified path do not exist in the file system (default: false)  

#### `channel.fromFilePairs`  
Emitting pair files (read1/read2).  
```{bash, eval = F}
Channel
  .fromFilePairs('/domus/h1/nimar/nextflow_tutorial/fastq/*{1,2}.fq')
  .view()
# [HG00100, [/domus/h1/nimar/nextflow_tutorial/fastq/HG00100_1.fq, /domus/h1/nimar/nextflow_tutorial/fastq/HG00100_2.fq]]
# [HG00097, [/domus/h1/nimar/nextflow_tutorial/fastq/HG00097_1.fq, /domus/h1/nimar/nextflow_tutorial/fastq/HG00097_2.fq]]
# [HG00101, [/domus/h1/nimar/nextflow_tutorial/fastq/HG00101_1.fq, /domus/h1/nimar/nextflow_tutorial/fastq/HG00101_2.fq]]
# 
Channel
  .fromFilePairs('/some/data/*', size: -1) { file -> file.extension }
  .view { ext, files -> "Files with the extension $ext are $files" }
# Files with the extension fq are [/domus/h1/nimar/nextflow_tutorial/fastq/HG00097_1.fq, /domus/h1/nimar/nextflow_tutorial/fastq/HG00097_2.fq, /domus/h1/nimar/nextflow_tutorial/fastq/HG00100_1.fq, /domus/h1/nimar/nextflow_tutorial/fastq/HG00100_2.fq, /domus/h1/nimar/nextflow_tutorial/fastq/HG00101_1.fq, /domus/h1/nimar/nextflow_tutorial/fastq/HG00101_2.fq]
```
- size `-1` means any file  
- The notation `->` in groovy is used to define a closure (`{}`) which is essentially an anonymous function (see below for further information).  
- `{ file -> file.extension }`: This is a closure that takes each file matched by the pattern and returns its extension. The file parameter represents each file, and file.extension gets its extension. This closure is used to group the files by their extensions.
- .view { ext, files -> ... }: This is another closure used to process and print each group of files. It takes two parameters:
  - ext: The file extension (the key by which files were grouped).  
  - files: The list of files that have this extension.  
  - println "Files with the extension $ext are $files": This prints out the extension and the corresponding files.  

#### `channel.fromSRA`  
Emitting files from SRA.  
```{bash, eval = F}
ids = ['ERR908507', 'ERR908506', 'ERR908505']
Channel
    .fromSRA(ids)
    .view()
```
Output:  
```
[SRR1448794, /vol1/fastq/SRR144/004/SRR1448794/SRR1448794.fastq.gz]
[SRR1448795, /vol1/fastq/SRR144/005/SRR1448795/SRR1448795.fastq.gz] 
[SRR1448796, /vol1/fastq/SRR144/006/SRR1448796/SRR1448796.fastq.gz]
```


#### `channel.of`  
Creating a channel from a list of values.  
```{bash, eval = F}
Channel
  .of( 'alpha', 'beta', 'delta' )
  .subscribe { println "Got: $it" }
```
Output:  
```
Got: alpha
Got: beta
Got: delta
```

or ranges:  
```{bash, eval = F}
Channel
  .of( 1..22, X..Y )
  .view()
```
Output:  
```
1  
2  
:  
22  
X  
Y  
```


#### `channel.topic` 
It is a channel type by which you can connect multiple processes.  
```{bash, eval = F}
process proc1 {
  val x, topic: 'my_topic'

  script:
  """
  echo "Processing $x"
  """
}

process proc2 {
  val x, topic: 'my_topic'

  script:
  """
  echo "Processing $x"
  """
}

channel
  .topic('my_topic')
  .view()
```



#### `channel.value`  
Creating a value channel by which you can bind the channel to a specific variable.  
```{bash, eval = F}
workflow_summary    = WorkflowSpatialxe.paramsSummaryMultiqc(workflow, summary_params)
ch_workflow_summary = Channel.value(workflow_summary)
```


#### `channel.watchPath`  
Check the presence of file with certain events which should be specified (`create`, `modify`, `delete`).  
```{bash, eval = F}
Channel
   .watchPath( '/path/*.fa', 'create, modify' )
   .subscribe { println "Fasta file: $it" }
```

**The watchPath factory waits endlessly for files that match the specified pattern and event(s), which means that it will cause your pipeline to run forever. Consider using the until operator to close the channel when a certain condition is met (e.g. receiving a file named DONE).**  






## Operators  
### `mix`  
emits the emited outputs of multiple channels.  
```{bash, eval = F}
c1 = Channel.from( 'alpha', 'beta', 'delta' )
c2 = Channel.from( 'one', 'two', 'three' ) 
c1.mix( c2 ).view()
```


## Workflow
### `take`  
Take is similar to input but when it is used the beignning of the workflow begins with `main`   
```{bash, eval = F}
workflow {
  take: 
    myChannel
 
  main:
    myProcess(myChannel)
}
```


# Misc
## `->`  
The notation `->` in groovy is used to define a closure (`{}`) which is essentially an anonymous function (see below for further information).  
Basic usage:  
```{bash, eval = F}
{param1, param2 -> 
  // code here
}
```
- `{}` body of the closure  
- `param1, param2` parameters of the closure
- `->` separates the parameters from the body of the closure.
Example:  
```{bash, eval = F}
def myClosure = {param1, param2 -> 
  println "param1: $param1"
  println "param2: $param2"
}
myClosure('hello', 'world')
```
Output:  
```
param1: hello
param2: world
```

You define a new variable in *groovy* by `String variable -> println "${variable}; ${variable.size()}".  
```{bash, eval = F}
Channel
    .from( 'alpha', 'beta', 'lambda' )
    .subscribe { String str ->
        println "Got: ${str}; len: ${str.size()}"
    }
```

## `LinkedHashMap`  
A `LinkedHashMap` is a collection that maintains the order of its elements. When you iterate over a `LinkedHashMap`, the elements are returned in the order they were inserted.  
```{bash, eval = F}
def create_fastq_channel(LinkedHashMap row) {
    // create meta map
    def meta = [:]
    meta.id         = row.sample
    meta.single_end = row.single_end.toBoolean()

    // add path(s) of the fastq file(s) to the meta map
    def fastq_meta = []
    if (!file(row.fastq_1).exists()) {
        exit 1, "ERROR: Please check input samplesheet -> Read 1 FastQ file does not exist!\n${row.fastq_1}"
    }
    if (meta.single_end) {
        fastq_meta = [ meta, [ file(row.fastq_1) ] ]
    } else {
        if (!file(row.fastq_2).exists()) {
            exit 1, "ERROR: Please check input samplesheet -> Read 2 FastQ file does not exist!\n${row.fastq_2}"
        }
        fastq_meta = [ meta, [ file(row.fastq_1), file(row.fastq_2) ] ]
    }
    return fastq_meta
}
``` 


# How to understand a nextflow pipeline?  
In general nextflow pipeline is orchestrated by **main.nf** file. This file is the entry point of the pipeline.  
and then **nextflow.config** file. This file contains the configuration of the pipeline.  
Suggested order of the file to look:  
1. main.nf: Understand the primary workflow structure.  
2. nextflow.config: Learn about configurations and profiles.  
3. modules: Explore the individual modules for different tasks.  
4. subworkflows: Check for any reusable sub-workflows used in the main workflow.  

## main.nf  

## Initialization and Parameter Definition:  

- Defines parameters and initial channels.  
- Specifies input files, directories, and any necessary configurations.  

```{bash, eval = F}
params.input = 'data/*'
params.outdir = 'results'

workflow.onComplete {
    println "Workflow completed"
}
```
- `params.input`: Specifies the input data files.  
- `params.outdir`: Defines the output directory.  
- `workflow.onComplete`: A block to execute when the workflow finishes.  

## Processes and Workflow Execution:  

- Contains process definitions that encapsulate individual computational steps.  
- Defines the workflow that connects these processes.  

```{bash, eval = F}
process process_data {
    input:
    path input_file from Channel.fromPath(params.input)
    
    output:
    path "${input_file.baseName}.processed"

    script:
    """
    process_tool -i $input_file -o ${input_file.baseName}.processed
    """
}
```  
- `process_data`: A process definition that takes an input file, processes it, and outputs the result.  
- `input_file`: Specify the input channel.    
- `output`: Defines the output path.  
- `script`: The shell command or script to run.  
- 
## Output and Reporting:  

- Handles the output of results and logging.  
- May include final steps for summarizing results and generating reports.  

```{bash, eval = F}
workflow {
    files = Channel.fromPath(params.input)
    processed_files = files | process_data
    
    processed_files.view {
        println "Processed file: $it"
    }
}
```

- `workflow`: Defines the workflow structure.  
- `files`: A channel created from the input files.  
- `processed_files`: The result of processing the files through `process_data`.  
- `processed_files.view`: Outputs the processed files for logging.    

## groovy class  
main.nf  may refer to some functions which can be defined in a groovy class.    
In Groovy, a class is a blueprint for creating objects, encapsulating data for the object, and methods to manipulate that data. Groovy classes are similar to Java classes but with some syntactic sugar that makes them easier to work with.  
```{bash, eval = F}
class Person {
    String name
    int age

    void sayHello() {
        println "Hello, my name is $name and I am $age years old."
    }
}

// Creating an instance of the class
def person = new Person(name: 'John', age: 30)
person.sayHello()
```
Key Points
- Properties: `String name` and `int age` are properties of the `Person` class.  
- Methods: `sayHello` is a method that prints a greeting message.  
- Instance Creation: `new Person(name: 'John', age: 30)` creates an instance of the `Person` class with specified properties.  
- Method Call: `person.sayHello()` calls the `sayHello` method on the person object.  

### `public`  
Access Modifier: Specifies that the member (field, method, or class) is accessible from any other class.  
```{bash, eval = F}
public class Person {
    public String name
    public int age

    public void sayHello() {
        println "Hello, my name is $name and I am $age years old."
    }
}
```
### `void`  
Return Type: Indicates that the method does not return any value.  
```{bash, eval = F}
void sayHello() {
    println "Hello"
}
```

### `static`  
Modifier: Means that the member belongs to the class, rather than instances of the class. Static members can be accessed without creating an instance of the class.  
```{bash, eval = F}
static String greet() {
    return "Hello"
}
``` 

